### 2016-11-25 Detector/Detector.tex
Bit more on the trigger

The tracker will be upgraded during the extended year-end technical stop in early 2017. %or just replaced?

@@

HCAL RESOLUTION!

\subsection{Muon system}
\label{sec:CMSLHC_CMS_muons}
One of the main physics goals of the \ac{CMS} experiment
at design was to search for $H\rightarrow ZZ\rightarrow 4\mu$ decays. Muons
suffer less from radiative losses in the tracker material, which means the most accurate
4-body mass can be constructed when searching for muon final states.
After discovery of the Higgs boson, the ZZ channel remains important
for Higgs measurements, while many \ac{BSM} searches make use 
they

A cross-section of one quadrant of the detector, indicating
the positions of the muon systems, is shown in figure \ref{fig:CMS_MuonSystem}.
\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.8\textwidth]{./Detector/Plots/MuonSystem.png}
\caption{Cross-section of one quadrant of the CMS detector, indicating
the positions of the different muon system components \cite{cms-tdr-vol1}.}
\label{fig:CMS_MuonSystem}
\end{center}
\end{figure}
%Because of the uncertainty in the eventual background rates and in the ability of the muon system to measure the correct beam-crossing time when the LHC reaches full luminosity, a com- plementary, dedicated trigger system consisting of resistive plate chambers (RPC) was added in both the barrel and endcap regions. The RPCs provide a fast, independent, and highly-segmented trigger with a sharp pT threshold over a large portion of the rapidity range (|η| < 1.6) of the muon system. The RPCs are double-gap chambers, operated in avalanche mode to ensure good operation at high rates. They produce a fast response, with good time resolution but coarser position reso- lution than the DTs or CSCs. They also help to resolve ambiguities in attempting to make tracks from multiple hits in a chamber.

\subsection{Triggering}
\label{sec:CMSLHC_CMS_trigger}

\subsection{Data processing}
\label{sec:CMSLHC_CMS_dataproc}

>>>>

The pixel tracker will be upgraded during the extended year-end technical stop in early 2017. %or just replaced?

@@

HCAL RESOLUTION!
%UPGRADE
During LS1, the \ac{HO} and \ac{HF} pmts were replaced

\subsection{Muon system}
\label{sec:CMSLHC_CMS_muons}
One of the main physics goals of the \ac{CMS} experiment
at design was to search for $H\rightarrow ZZ\rightarrow 4\mu$ decays. Muons
suffer less from radiative losses in the tracker material, which means the most accurate
4-body mass can be constructed when using the dimuon final state of the $Z$'s.
Since the discovery of the Higgs boson the ZZ channel has continued to be an important
final state for Higgs measurements, at the same time many \ac{BSM} searches also make use of 
signatures with muons in the final state.
All of this means that good muon identification efficiency and precise muon momentum resolution
are very important. The \ac{CMS} muon system is made up out of three different subsystems which
are all necessary to ensure the identification requirements can be met. It is located
outside of the magnet coil, interspersed in the iron return yoke of the solenoid.
In the barrel of the detector the \ac{DT} chambers provide coverage for $|\eta|<1.2$, with
the \ac{CSC} covering the $1.2<|\eta|<2.4$ region. The \ac{RPC}, covering the same
regions as both \ac{DT} and \ac{CSC}, is used as a dedicated triggering system.

A cross-section of one quadrant of the detector, indicating
the positions of the \ac{DT},\ac{CSC} and \ac{RPC}, is shown in figure \ref{fig:CMS_MuonSystem}. 
\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.8\textwidth]{./Detector/Plots/MuonSystemUpgrade.png}
\caption{Cross-section of one quadrant of the CMS detector, indicating
the positions of the different muon system components \cite{cms-muon-upgrade}. The 
fourth muon endcap stations were extended during \ac{LS1}, in Run 1 only the lower
part existed.}
\label{fig:CMS_MuonSystem}
\end{center}
\end{figure}

The \ac{DT} consists of

%UPGRADE
During LS1, the endcaps of the muon system were upgraded. Each endcap received an additional muon station, extending ME/4 (figure \ref{fig:CMS_MuonSystem}) out towards
$R\approx 700 $cm, in line with the other three muon stations \cite{cms-muon-upgrade}.
%Because of the uncertainty in the eventual background rates and in the ability of the muon system to measure the correct beam-crossing time when the LHC reaches full luminosity, a com- plementary, dedicated trigger system consisting of resistive plate chambers (RPC) was added in both the barrel and endcap regions. The RPCs provide a fast, independent, and highly-segmented trigger with a sharp pT threshold over a large portion of the rapidity range (|η| < 1.6) of the muon system. The RPCs are double-gap chambers, operated in avalanche mode to ensure good operation at high rates. They produce a fast response, with good time resolution but coarser position reso- lution than the DTs or CSCs. They also help to resolve ambiguities in attempting to make tracks from multiple hits in a chamber.

\subsection{Triggering and data processing}
\label{sec:CMSLHC_CMS_trigger}
When the \ac{LHC} is operating under standard conditions, protons are collided at a rate
of 40 MHz. This rate is too high for the \ac{DAQ} system to read out every event. In
addition this high collision rate produces such a large number of events that it would
not be feasible to write every single event, with sizes of around 1 MB, to tape. A trigger
system is therefore used to select events of interest for study, which reduces the rate
at which events are written to disk to  $\mathcal{O}(1 \text{ kHz})$.

The trigger system consists of two stages: the \ac{L1} trigger, which only makes use of
front-end electronics 
and reduces the event rate from 40 MHz to around 100 kHz, and the
\ac{HLT}, which runs on a computer farm with around 13000 CPU cores \cite{cms-trigger}.
During \ac{LS1} and the year-end technical stop at the end of 2015 the \ac{L1} trigger was
replaced by a completely upgraded system, which provides amongst others 
improved object isolation, hadronic tau identification, muon \pT resolution and
jet finding.

An overview of the data flow in the \ac{L1} trigger is given in figure \ref{fig:CMS_Trigger}.
The \ac{L1} trigger only makes use of information from the calorimeters and the muon systems.
The calorimeter trigger starts from \ac{HCAL} and \ac{ECAL} energy deposits, 
which are fed to the first layer of the trigger. This layer maps onto
columns of the detector, receiving data from many bunch crossings. These data are passed
on to the second layer of the trigger in a way that bundles data from a single bunch crossing, but
different columns of the detector, to be processed on one node in the second layer of the
trigger. In this step basic object identification is performed.

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.5\textwidth]{./Detector/Plots/CMSTrigger.png}
\caption{Schematic of the different components of the \ac{L1} trigger \cite{cms-trigger-upgrade}.}
\label{fig:CMS_Trigger}
\end{center}
\end{figure}
